\documentclass{article}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{hyperref}

\usepackage{amssymb}

\usepackage{amsmath}  

\usepackage{algorithmic}
\usepackage{algorithm}  
% \usepackage{algpseudocode}  
\usepackage{amsmath}  
\renewcommand{\algorithmicrequire}{\textbf{Input:}}  % Use Input in the format of Algorithm  
\renewcommand{\algorithmicensure}{\textbf{Output:}} % Use Output in the format of Algorithm    
\usepackage{listings}
\usepackage{url}

\usepackage{etoolbox}
\newtoggle{solution}
\toggletrue{solution}
%\togglefalse{solution}
\usepackage{color}

\usepackage{listings}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=Python,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

\newcommand{\solution}[2][0pt]{\iftoggle{solution}{\smallskip{\color{red}{\flushleft\textbf{Solution}:}\par#2}}{\vspace*{#1}}}
\renewcommand{\baselinestretch}{1.2}%Adjust Line Spacing
%\geometry{left=2.0cm,right=2.0cm,top=2.0cm,bottom=2.0cm}% Adjust Margins of the File

% Create horizontal rule command with an argument of height
\newcommand{\horrule}[1]{\rule{\linewidth}{#1}}
% Set the title here
\title{
    \normalfont \normalsize
    \textsc{ShanghaiTech University} \\ [25pt]
    \horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
    \huge CS240 Algorithm Design and Analysis \\ % The assignment title
    \LARGE Fall 2023\\
    \LARGE Problem Set 3\\
    \horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}
\author{}
\date{Due: 23:59, Dec. 12, 2023}

% Add the support for auto numbering
% use \problem{title} or \problem[number]{title} to add a new problem
% also \subproblem is supported, just use it like \subsection
\newcounter{ProblemCounter}
\newcounter{oldvalue}
\newcommand{\problem}[2][-1]{
	\setcounter{oldvalue}{\value{secnumdepth}}
	\setcounter{secnumdepth}{0}
	\ifnum#1>0
		\setcounter{ProblemCounter}{#1}
	\else
		\stepcounter{ProblemCounter}
	\fi
	\section{Problem \arabic{ProblemCounter}: #2}
	\setcounter{secnumdepth}{\value{oldvalue}}
}
\newcommand{\subproblem}[1]{
	\setcounter{oldvalue}{\value{section}}
	\setcounter{section}{\value{ProblemCounter}}
	\subsection{#1}
	\setcounter{section}{\value{oldvalue}}
}

\begin{document}
\maketitle
\vspace{3ex}

\begin{enumerate}
%\item Please write your solutions in English. 
\item Submit your solutions to Gradescope (www.gradescope.com).
\item In ``Account Settings'' of Gradescope, set your FULL NAME to your Chinese name and enter your STUDENT ID correctly. 
\item If you want to submit a handwritten version, scan it clearly. Camscanner is recommended. 
\item When submitting your homework, match each of your solution to the corresponding problem number. 
%\item Violations to any of above may result in some penalty on your score. 
\end{enumerate}

\newpage
\problem{}
\textbf{NAE-4-SAT}: A clause in an instance of $4$-SAT is a disjunction with four literals i.e. $(x_1 \vee x_2 \vee x_3 \vee x_4)$. A satisfied assignment for a collection of clauses is called \emph{not all equal} (NAE) if the literals in each clause are not all equal to each other. In other words, at least one false, and at least one true. Prove NAE-4-SAT is NP-complete, using a reduction from 3-SAT.


\newpage
\problem{}
\textbf{NAE-3-SAT}: Prove NAE-3-SAT is NP-complete.


\newpage
\problem{}
\textbf{4-SAT}: Prove 4-SAT is NP-complete.


\newpage
\problem{}
\textbf{Feedback Vertex Set problem} Let $G = (V, E)$ be a directed graph. A set $F \subseteq V$ is a \emph{feedback vertex set} if every cycle of $G$ contains at least one vertex from $F$. The \emph{Feedback Vertex Set problem} asks whether $G$ has a feedback vertex set with at most $K$ vertices.  Show that Feedback Vertex Set is
NP-complete.  (Hint:  use
Vertex Cover problem for reduction.)


\subsubsection*{Vertex Cover Problem}
Given an undirected graph $G' = (V', E')$ and a number $K'$, the Vertex Cover problem asks if there is a vertex cover of size at most $K'$, i.e., a subset of vertices $V'' \subseteq V'$ such that every edge in $E'$ is incident to at least one vertex in $V''$.


\newpage
\problem{}
\textbf{STINGY SAT} is the following problem: given a set of clauses (each a disjunction of literals) and an integer k, find a satisfying assignment in which at most k variables a true, if such an assignment exists. Prove that STINGY SAT is NP-complete.


\newpage

\problem{}
Given a tree $T$, a set of terminal vertices, and an integer $k$, is there a set of at most $k$ edges which, when removed from $T$, separates every pair of terminal vertices? Prove that this problem is NP-complete.








\end{document}
